---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
library(magrittr)
devtools::load_all()
```

# hddtest

Functions for hypothesis testing of high dimensional discrete data. Currently functions for multinomial data, as described in [1], are available. Functions for multivariate binary data will be added in the future.  

## Installation

You can install hddtest from github with:

```{r gh-installation, eval = FALSE}
install.packages("devtools")
devtools::install_github("AmandaRP/hddtest")
library("hddtest")
```

## Example

Generate two multinomial count vectors and test whether they come from the same underlying distribution:

```{r example}
data <- genMultinomialData(null_hyp=FALSE,sample_size = 1)
multinom.test(x=data[[1]],y=data[[2]])
```

The last call can also be done in a "tidy" fashion using the forward pipe from the `magrittr` package:

```{r}
data %>% multinom.test
```


## References

[1] Plunkett, A. & Park, J. (2018) Two-sample test for sparse high-dimensional multinomial distributions, TEST, doi.org/10.1007/s11749-018-0600-8

[2] Plunkett, A. & Park, J. (2017) Two-sample tests for sparse high-dimensional binary data, Communications in Statistics - Theory and Methods, 46:22, 11181-11193, DOI: 10.1080/03610926.2016.1260743
